Rank Verification for Exponential Families

### 2016-08-05 verifying-winner.tex
broken sentences fixed; unified notations

\begin{enumerate}
\item {\em Is Trump really winning?} If Trump and Cruz were in fact tied, then Trump's share of their combined 490 votes would be distributed as $\text{Binom}\left(490, 0.5\right)$. Because the (two-tailed) $p$-value for this pairwise test is $p = 0.006$, we reject the null and conclude that Trump is really winning.

@@

Previous works construct tests comparing parameters in various families of distributions, by comparing just the winner and the runner-up. \citet{Gutmann:1987fk} tackles the question where the observations are from a location family, and the location parameters are compared. In which case, the difference of the winner and the runner-up has to pass a predetermined threshold. In contrast, a related paper by \citet{Karnnan:2009iv} discusses a scale family, comparing the scale parameters instead. Specifically, to test for the smallest variance across multiple populations, it is sufficient to test the ratio of the smallest to the second-smallest variances. \citet{Bofinger:1991hv} gives a location-scale family generated by translating an exponential distribution as well. Similarly, test the difference between the winner and the runner-up is enough. A follow-up paper by \citet{Maymin:1992fz} gave an analogous result for location-scale families, yet still under the constraints of independence as well.

However, as location-scale families have limited overlap with exponential families, the counterpart question for exponential families remains unaddressed. Location-scale families assumes independence between observations, while this is not required in exponential families. For instance, in the vicinity of the polling problem above, \citet{Nettleton:2009ht} provides an asymptotic test for multinomial distribution, imposing the selection by considering intersection-union tests. \citet{Ng:2007cn} also analyzes the multinomial case for an exact test, but with an additional restriction that maximum observation is a known constant $M$. In terms of the polling problem above, we poll until one candidate hits a goal $M$. While in either papers the observations $X$ are not independent, it still suffices to test the difference between the winner and the runner-up.

One final result that our particular interest in the subset selection problem \citep{Gupta:1967wg}. It can indeed be used to solve the multinomial case. They select a subset that contains the largest, or ``tagged'', parameter with probability $1-\alpha$. Confirming the largest parameter is thus equivalent to asserting the selected subset has size $1$. \Cref{sec:subsetsel} includes a comparison of our proposed test and their test.

@@

To avoid notational clutter, we consider only the case where the winner is $1$ and the runner-up is $2$, etc., i.e.\ $X_1 \ge \cdots \ge X_n$. For other cases, the following proof remains valid under relabeling. In this case, we will test the null hypothesis $H_{01}: \theta_1 \le \max_{j > 1} \theta_j$, which is the union of the null hypotheses $H_{01j}: \theta_1 \le \theta_j$ for $j \ge 2$. For each of these we can construct an exact $p$-value $p_{1j}$, which is valid under $H_{01j}$ conditional on $A_1$, the event that $X_1$ is the winner. Hence by \Cref{lma:union}, a test based on $p_{1*} = \max_j p_{1j}$ would be valid for $H_{01}$ conditional on $A_1$. Procedure 1 performs an unadjusted pairwise test comparing $X_1$ to $X_2$. Hence it is sufficient to show that $p_{12} = p_{1*}$ and that the test based on $p_{12}$ coincides with the unadjusted pairwise test.

@@

Now suppose the exponential family is defined on the counting measure instead. If ties are broken independently and randomly, the end points on the rays can be considered as ``half an atom'' if the coordinates are integers, and a ``full atom'' otherwise. The number of atoms on each ray is the same (after the extension $\tilde{A}$) and thus the atoms on each ray can be paired up. The inequalities above still hold for each pair of the atoms.
\end{proof}

@@

\begin{proof}
This can be fitted into the selective sequential model selection framework proposed by \citet{Fithian:2015uj}. Given the observations $X_1 \ge \cdots \ge X_n$, drawn from an unknown distribution $F$, we consider the following sequence of null hypotheses:
$$H_{0j}: \theta_j \le \max_{k > j} \theta_k.$$
Alternatively, we can think of them as a sequence of nested model
$$\mathcal{M}_0\left(X\right) \subseteq \mathcal{M}_1\left(X\right) \subseteq \cdots \subseteq \mathcal{M}_n\left(X\right), ~~~~~~~~ \mathcal{M}_j\left(X\right)^c = \left\{\theta_1 > \cdots > \theta_j > \max_{k > j} \theta_k\right\},$$
and the each $\mathcal{M}_j$ depends on $X$ through the order of $X$. The step-down procedure for rejecting $H_{0j}$ is equivalent to selecting the smallest $j$ for which we cannot reject $F \in \mathcal{M}_j\left(X\right)$.

@@

Hence similar to the proof of Part 1 of \Cref{thm:main}, Schur-concavity ensures $p_{j\left(j+1\right)} \ge p_{jk}$ for all $k>j$, meaning that it is sufficient to compare $X_j$ and $X_{j+1}$. In fact, we can take $p_j = p_{j\left(j+1\right)}$. These are valid selective $p$-value as required in \citet{Fithian:2015uj}, as
$$\PP_\theta \left[p_j \le \alpha \mid \theta_{[k]} \text{ is not the } k\text{-th largest, for some } k \le j\right] \le \alpha \text{ a.s.}, ~~~~~~~~ \text{for all } \theta,$$
by tower property. Note that the conditioning is indeed a random event as the order of $\theta$ depends on $X$ (emphasized with the square brackets here). \KHcomment{I should probably add a line transitioning from using $\mathcal{M}$ to suddenly mentioning $\theta$ here, shouldn't I?}

>>>>

\begin{enumerate}
\item {\em Is Trump really winning?} If Trump and Cruz were in fact tied, then Trump's share of their combined 490 votes would be distributed as $\text{Binomial}\left(490, 0.5\right)$. Because the (two-tailed) $p$-value for this pairwise test is $p = 0.006$, we reject the null and conclude that Trump is really winning.

@@

Previous works construct tests comparing parameters in various families of distributions, by comparing just the winner and the runner-up. \citet{Gutmann:1987fk} tackles the question where the observations are from a location family, and the location parameters are compared. In which case, the difference of the winner and the runner-up has to pass a predetermined threshold. In contrast, a related paper by \citet{Karnnan:2009iv} discusses a scale family, comparing the scale parameters instead. Specifically, to test for the smallest variance across multiple populations, it is sufficient to test the ratio of the smallest to the second-smallest variances. \citet{Bofinger:1991hv} gives a location-scale family generated by translating an exponential distribution as well. Similarly, test the difference between the winner and the runner-up is enough. A follow-up paper by \citet{Maymin:1992fz} gave an analogous result for location-scale families, yet still under the constraints of independence.

However, as location-scale families have limited overlap with exponential families, the counterpart question for exponential families remains unaddressed. Location-scale families assumes independence between observations, while this is not necessarily the case in exponential families. For instance, in the vicinity of the polling problem above, \citet{Nettleton:2009ht} provides an asymptotic test for multinomial distribution, imposing the selection by considering intersection-union tests. \citet{Ng:2007cn} also analyzes the multinomial case for an exact test, but with an additional restriction that maximum observation is a known constant $M$. In terms of the polling problem above, we poll until one candidate hits a goal $M$. While in either papers the observations $X$ are not independent, it still suffices to test the difference between the winner and the runner-up.

One final result that has our attention is the subset selection problem \citep{Gupta:1967wg}. The procedure provided in the paper can indeed be used to solve the multinomial case. It selects a subset that contains the largest, or ``tagged'', parameter with probability $1-\alpha$. Confirming the largest parameter is thus equivalent to asserting the selected subset has size $1$. \Cref{sec:subsetsel} includes a comparison of our proposed test and their test.

@@

To avoid notational clutter, we consider only the case where the winner is $1$ and the runner-up is $2$, etc., i.e.\ $X_1 \ge \cdots \ge X_n$. For other cases, the following proof remains valid under relabeling and can thus be applied. In this case, we will test the null hypothesis $H_{01}: \theta_1 \le \max_{j > 1} \theta_j$, which is the union of the null hypotheses $H_{01j}: \theta_1 \le \theta_j$ for $j \ge 2$. For each of these we can construct an exact $p$-value $p_{1j}$, which is valid under $H_{01j}$ conditional on $A_1$, the event that $X_1$ is the winner. Hence by \Cref{lma:union}, a test based on $p_{1*} = \max_j p_{1j}$ would be valid for $H_{01}$ conditional on $A_1$. Procedure 1 performs an unadjusted pairwise test comparing $X_1$ to $X_2$. Hence it is sufficient to show that $p_{12} = p_{1*}$ and that the test based on $p_{12}$ coincides with the unadjusted pairwise test.

@@

Now suppose the exponential family is defined on the counting measure instead. If ties are broken independently and randomly, the end points on the rays can be considered as ``half an atom'' if the coordinates are integers, and a ``full atom'' otherwise. The number of atoms on each ray is the same (after the extension $\tilde{A}$) and thus the atoms on each ray can be paired up. The inequalities above still hold for each pair of the atoms. Summing these inequalities yields our desired result.
\end{proof}

@@

\begin{proof}
This can be fitted into the selective sequential model selection framework proposed by \citet{Fithian:2015uj}. Given the observations $X_1 \ge \cdots \ge X_n$, drawn from an unknown distribution $F$, we consider the following sequence of null hypotheses:
$$H_{0j}: \theta_j \le \max_{k > j} \theta_k.$$
Alternatively, we can think of them as a sequence of nested model
$$\mathcal{M}_0\left(X\right) \subseteq \mathcal{M}_1\left(X\right) \subseteq \cdots \subseteq \mathcal{M}_n\left(X\right), ~~~~~~~~ \mathcal{M}_j\left(X\right)^c = \left\{\theta_{[1]} > \cdots > \theta_{[j]} > \max_{k > j} \theta_{[k]}\right\},$$
and the each $\mathcal{M}_j$ depends on $X$ through the order of $X$. The step-down procedure for rejecting $H_{0j}$ is equivalent to selecting the smallest $j$ for which we cannot reject $F \in \mathcal{M}_j\left(X\right)$.

@@

Hence similar to the proof of Part 1 of \Cref{thm:main}, Schur-concavity ensures $p_{j\left(j+1\right)} \ge p_{jk}$ for all $k>j$, meaning that it is sufficient to compare $X_j$ and $X_{j+1}$. In fact, we can take $p_j = p_{j\left(j+1\right)}$. These are valid selective $p$-value as required in \citet{Fithian:2015uj}, as
\begin{align*}
\PP_\theta \left[p_j \le \alpha \mid \theta_{[k]} \text{ is not the } k\text{-th largest, for some } k \le j\right] = \PP_\theta \left[p_j \le \alpha \mid \mathcal{M}_j\left(X\right)^c = M\right] \le \alpha \text{ a.s.},
\end{align*}
for all $\theta \in M$, where $M$ is the selected model, by tower property. Note that the conditioning is indeed a random event as the order of $\theta$ depends on $X$ (emphasized with the square brackets here).
